Astr 513: Final Project

### 2016-05-08 Final_Project/ms.tex
Changed little things in labels and captions, rearranged plots

\documentclass[12pt,a4paper]{article}
\usepackage{amsmath}
\usepackage[english]{babel}
\usepackage{graphicx}
\usepackage{listings}
\usepackage{fullpage}
\usepackage[T1]{fontenc}
\usepackage{enumerate}
\usepackage[makeroom]{cancel}
\usepackage{hyperref}
\usepackage{natbib}
\bibliographystyle{apj}

@@

The correlation matrix is calculated from the covariance matrix according to
%
\begin{equation}
  \label{eq:correlation}
  Corr_{i,j} = \frac{\mathbf{C}_{i,j}}{\sqrt{\mathbf{C}_{i,i}}\sqrt{\mathbf{C}_{j,j}}}.
\end{equation}
%
Note that the correlation of a variable with itself is 1 by definition.
The correlation matrices for both of our data sets can be seen below in
\autoref{fig:cmatrix_both}.
%
\begin{figure}
  \includegraphics[width=0.5\linewidth]{figures/Betoule_correlation.eps}
  \includegraphics[width=0.5\linewidth]{figures/Rest_correlation.eps}
  \caption{Correlation matrix of $\mu_0(z)$ from the
    \citet{betoule2014} sample on the left and \citet{rest2014} sample 
    on the right. Very few correlations exist in \citet{betoule2014} likely 
    due to the averaging over redshift bins. Significant correlation exists in 
    \citet{rest2014} due to dependencies on systematic variables
    as seen in \autoref{eq:RestCovMatrix}.}
%  \label{fig:cmatrix_betoule}
%\end{figure}
%
%\begin{figure}
%  \includegraphics[width=0.8\linewidth]{figures/Rest_correlation.eps}
%  \caption{Correlation matrix of $\mu_0(z)$ from the
%    \citet{rest2014} sample. Significant correlation exists between
%    different samples due to dependencies on systematic variables
%    as seen in \autoref{eq:RestCovMatrix}.}
  \label{fig:cmatrix_both}
\end{figure}

@@

\subsection{Likelihood}
\label{sec:likelihood}
Using the above model for the distance modulus allows us to write a
probability that a model can generate our data:
%
\begin{align}
  \label{eq:likelihood}
  P(\mu(z)|\mathrm{cosmo}) &= \frac{1}{\sqrt{2\pi|\mathbf{C}|}}\nonumber \\
  &\qquad{}\exp\left[-\frac{1}{2} \sum_{i,j=0}^{N_{samples}}\left(\mu(z) - \mu_{\mathrm{model}}(z)\right)_i 
    \mathbf{C}^{-1}_{i,j}\left(\mu(z) - \mu_{\mathrm{model}}(z)\right)_j\right]
\end{align}
%
where $\mathbf{C}$ is the full covariance matrix between all of the $N_{samples}$ 
in the data, and $\mathrm{cosmo}$ represents the cosmology that we are investigating
(e.g. flat $\Lambda$CDM, $w$CDM, etc.). 
This is not the full picture though, since we must employ
Bayes' Theorum in order to include any priors on our choice of 
parameters and recover the posterior distribution function of our
parameters given our data
%
\begin{equation}
  \label{eq:posterior}
  P(\mathrm{cosmo}|\mu(z)) = P_{prior}(\mathrm{cosmo}) P(\mu(z)|\mathrm{cosmo}).
\end{equation}
%
For this work we investigate a combination of priors. \CosmoSIS is implemented
in a way that requires the user to provide bounds on the parameter space, which we
detail for each parameter in \autoref{tab:priors}
Within those bounds, we investigated using a flat prior $P_{prior}=1$ and
\Planck priors.

@@

\subsection{Planck Priors}
\label{sec:planck_priors}
Our priors are given by the analysis of the \citet{planck2013} data set.
However, instead of simply using the posterior mean and variance of
the parameters from that analysis as priors in our project, we
employed the full \Planck likelihood as a prior. That is,
we first run the Code for Anisotropies in the Microwave Background (CAMB)
in order to obtain the temperature power spectrum $\mathcal{D}_l^{TT}$
and the polarization power spectrum $\mathcal{C}_l^{EE}$
and feed it into the \Planck likelihood code\footnote{More information
on the \Planck likelihood can be found here: 
\url{https://wiki.cosmos.esa.int/planckpla2015/index.php/CMB_spectrum_\%26_Likelihood_Code}.}.
The resulting likelihood
is our prior, and it captures any degeneracies in parameter space
observed by the \Planck analysis, as seen in \autoref{fig:planck_corner}. 
%
\begin{figure}
  \centering
  \includegraphics[width=0.5\linewidth]{figures/corner_example.eps}
  \caption{Posterior distributions of the cosmological parameter
    analysis using just the \Planck likelihood. Contours denote the
    68\% and 95\% confidence projected areas; dashed lines in
    the 1D histograms refer to the 68\% confidence intervals.}
  \label{fig:planck_corner}
\TM{THIS IS A PLACE HOLDER PLOT!} 
\end{figure}
%

@@

\section{Results}
\label{sec:results}
%\LC{Not sure where to put why we couldn't do the full analysis of Rest's data. 
%  I'm putting it here for now.}
%A complete analysis using the data from \citet{rest2014} is not done 
%in our paper since there are several issues with the publicly released 
%data. It is not apparent which uncertainties belonged to the statistical 
%covariance matrix since only the systematic covariance matrix is clearly 
%presented. Moreover, intrinsic dispersion of the data and photometric errors 
%of the distances depend on nuisance parameters in addition to the covariances 
%between the fit parameters. Fine-tuning nuisance parameters places a significant 
%time restraint on our work and is beyond the scope of this paper. For these reasons, 
%we only present the results for \citet{rest2014} that do not include the priors 
%from \citet{planck2013}.

@@

@@

  \includegraphics[width=0.5\linewidth]{figures/betoule_wcdm_withprior_bestfit.eps}
  \includegraphics[width=0.5\linewidth]{figures/rest_wcdm_withprior_bestfit.eps}
  \caption{Best fit model of the \citet{betoule2014} data (left) and \citet{rest2014} data (right) including the
    \Planck priors for a $w$CDM model. Blue points are the data with
    error bars corresponding to the diagonal of the provided covariance martrix.
    The solid black line is the most likely model from our MCMC chain.}

@@

    analysis of the \citet{betoule2014} data for a $w$CDM model. The left is without 
    \Planck priors and the right is with \Planck priors. Contours denote the
    68\% and 95\% confidence projected areas; dashed lines in
    the 1D histograms refer to the 68\% confidence intervals.}

@@

  \caption{Same as ~\autoref{fig:betoule_wcdm_corner} but with the \citet{rest2014} data.}

@@

\EW{Wasn't sure where to put the \Planck only corner plot}

@@

  \includegraphics[width=0.5\linewidth]{figures/nosn_wcdm_corner.eps}
  \caption{Same as ~\autoref{fig:betoule_wcdm_corner} and ~\autoref{fig:rest_wcdm_corner}, but using only \Planck likelihoods without \sn{} data.}
  \label{fig:nosnia_wcdm_corner}

@@

    analysis of the \citet{betoule2014} data for a $\Lambda$CDM model. The left is without 
    \Planck priors and the right is with \Planck priors. Contours denote the
    68\% and 95\% confidence projected areas; dashed lines in
    the 1D histograms refer to the 68\% confidence intervals.}

>>>>

\documentclass[12pt,a4paper]{article}
\usepackage{amsmath}
\usepackage[english]{babel}
\usepackage{graphicx}
\usepackage{listings}
\usepackage{fullpage}
\usepackage[T1]{fontenc}
\usepackage{enumerate}
\usepackage[makeroom]{cancel}
\usepackage{hyperref}
\usepackage{natbib}
\bibliographystyle{apj}
\usepackage[labelfont=bf]{caption}
\usepackage[labelsep=period]{caption}

@@

The correlation matrix is calculated from the covariance matrix according to
%
\begin{equation}
  \label{eq:correlation}
  Corr_{i,j} = \frac{\mathbf{C}_{i,j}}{\sqrt{\mathbf{C}_{i,i}}\sqrt{\mathbf{C}_{j,j}}}.
\end{equation}
%
Note that the correlation of a variable with itself is 1 by definition.
The correlation matrices for both of our data sets can be seen below in
\autoref{fig:cmatrix_both}.
%
\begin{figure}
  \includegraphics[width=0.5\linewidth]{figures/Betoule_correlation.eps}
  \includegraphics[width=0.5\linewidth]{figures/Rest_correlation.eps}
  \caption{{\it Left}: Correlation matrix of $\mu_0(z)$ from the
    \citet{betoule2014} sample. {\it Right}: Correlation matrix of $\mu_0(z)$ from 
	the \citet{rest2014} sample. Very few correlations exist in 
        \citet{betoule2014} likely due to the averaging over redshift bins. 
	Significant correlation exists in \citet{rest2014} due to dependencies 
	on systematic variables as seen in \autoref{eq:RestCovMatrix}.}
  \label{fig:cmatrix_both}
\end{figure}

@@

\subsection{Likelihood}
\label{sec:likelihood}
Using the above model for the distance modulus allows us to write a
probability that a model can generate our data:
%
\begin{align}
  \label{eq:likelihood}
  P(\mu(z)|\mathrm{cosmo}) &= \frac{1}{\sqrt{2\pi|\mathbf{C}|}}\nonumber \\
  &\qquad{}\exp\left[-\frac{1}{2} \sum_{i,j=0}^{N_{samples}}\left(\mu(z) - \mu_{\mathrm{model}}(z)\right)_i 
    \mathbf{C}^{-1}_{i,j}\left(\mu(z) - \mu_{\mathrm{model}}(z)\right)_j\right]
\end{align}
%
where $\mathbf{C}$ is the full covariance matrix between all of the $N_{samples}$ 
in the data, and $\mathrm{cosmo}$ represents the cosmology that we are investigating
(e.g. flat $\Lambda$CDM, $w$CDM, etc.). 
This is not the full picture though, since we must employ
Bayes' Theorum in order to include any priors on our choice of 
parameters and recover the posterior distribution function of our
parameters given our data
%
\begin{equation}
  \label{eq:posterior}
  P(\mathrm{cosmo}|\mu(z)) = P_{prior}(\mathrm{cosmo}) P(\mu(z)|\mathrm{cosmo}).
\end{equation}
%
For this work we investigate a combination of priors. \CosmoSIS is implemented
in a way that requires the user to provide bounds on the parameter space, which we
detail for each parameter in \autoref{tab:priors}.
Within those bounds, we investigated using a flat prior $P_{prior}=1$ and
\Planck priors.

@@

\subsection{Planck Priors}
\label{sec:planck_priors}
Our priors are given by the analysis of the \citet{planck2013} data set.
However, instead of simply using the posterior mean and variance of
the parameters from that analysis as priors in our project, we
employed the full \Planck likelihood as a prior. That is,
we first run the Code for Anisotropies in the Microwave Background (CAMB)
in order to obtain the temperature power spectrum $\mathcal{D}_l^{TT}$
and the polarization power spectrum $\mathcal{C}_l^{EE}$
and feed it into the \Planck likelihood code\footnote{More information
on the \Planck likelihood can be found here: 
\url{https://wiki.cosmos.esa.int/planckpla2015/index.php/CMB_spectrum_\%26_Likelihood_Code}.}.
The resulting likelihood
is our prior, and it captures any degeneracies in parameter space
observed by the \Planck analysis, as seen in \autoref{fig:planck_corner}. 
%
\begin{figure}
  \centering
  \includegraphics[width=0.5\linewidth]{figures/nosn_wcdm_corner.eps}
  \caption{Posterior distributions of the cosmological parameter
    analysis using just the \Planck likelihood. Contours denote the
    68\% and 95\% confidence projected areas; dashed lines in
    the 1D histograms refer to the 68\% confidence intervals.}
  \label{fig:planck_corner}
\end{figure}
%

@@

\section{Results}
\label{sec:results}
%\LC{Not sure where to put why we couldn't do the full analysis of Rest's data. 
%  I'm putting it here for now.}
%A complete analysis using the data from \citet{rest2014} is not done 
%in our paper since there are several issues with the publicly released 
%data. It is not apparent which uncertainties belonged to the statistical 
%covariance matrix since only the systematic covariance matrix is clearly 
%presented. Moreover, intrinsic dispersion of the data and photometric errors 
%of the distances depend on nuisance parameters in addition to the covariances 
%between the fit parameters. Fine-tuning nuisance parameters places a significant 
%time restraint on our work and is beyond the scope of this paper. For these reasons, 
%we only present the results for \citet{rest2014} that do not include the priors 
%from \citet{planck2013}. 

@@

    analysis of the \citet{betoule2014} data for a $\Lambda$CDM model. Contours denote the
    68\% and 95\% confidence projected areas and dashed lines in
    the 1D histograms refer to the 68\% confidence intervals. {\it Left}: without 
    \Planck priors. {\it Right}: with \Planck priors.}

@@

@@

    analysis of the \citet{betoule2014} data for a $w$CDM model. Contours denote the
    68\% and 95\% confidence projected areas and dashed lines in
    the 1D histograms refer to the 68\% confidence intervals. {\it Left}: without 
    \Planck priors. {\it Right}: with \Planck priors.}

@@

    analysis of the \citet{rest2014} data for a $w$CDM model. Contours denote the
    68\% and 95\% confidence projected areas and dashed lines in
    the 1D histograms refer to the 68\% confidence intervals. {\it Left}: without 
    \Planck priors. {\it Right}: with \Planck priors.}

@@

@@

  \includegraphics[width=0.48\linewidth]{figures/betoule_wcdm_withprior_bestfit.eps}
  \includegraphics[width=0.48\linewidth]{figures/rest_wcdm_withprior_bestfit.eps}
  \caption{{\it Left}: best fit model of the \citet{betoule2014} data for a $w$CDM model 
     with \Planck priors. {\it Right}: best fit model of the \citet{rest2014} data for a $w$CDM model 
     with \Planck priors. Blue points are the data with error bars corresponding 
     to the diagonal of the provided covariance martrix. The solid black line is 
     the most likely model from our MCMC chain.}

@@
